---
title: "P&A hand priming demo analysis"
author: "Cordula Vesper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(signal)
setwd("~/Desktop/Models of Perception and Action Exam")
```

Part 1. Remind yourselves what we did in the study and what we expected and why.

Simple implementation of a hand-priming-effect with 4 conditions (distractors/target): 1) high/low; 2) low/low; 3) high/high; 4) low/high


Part 2. Load one trial of your choice into R. The four columns refer to: sample, condition, x and y.

```{r}
participant <- 21
trial <- 12
filename <- paste(participant,'/',participant,'_',trial,'.txt',sep="")
hp = read.table(filename,header=FALSE,sep=",",dec=".")
names(hp) <- c("sample","cond","x","y") 
```

Part 3. Plot x against y.

```{r}
plot(hp$x,hp$y,'l')
```

Part 4. Map the y data to a more intuitive screen output with 0/0 in the lower left corner and where upwards movements are increasingly positive (i.e. so that it looks like the movements we made on the screen). 

This was unnecessary, sorry!


Part 5. Calculate the distance travelled (for x and y combined).
# No pythagoras - don't worry! 
# diff = default lag1, which means that it takes the previous value and calculates the difference. 
# Raised to the power of two because we don't want the negative and positive values to cancel each other out. Then sqrt.
```{r}
dist <- sum(sqrt(diff(hp$x)^2 + diff(hp$y)^2)) 
paste("Distance: ",round(dist))
# Can also just use print
print(dist)
```
# distance is 6387 pixels. 


Part 6. Add a time vector. I forgot to measure time so I had to estimate it post-hoc (fine for this exercise, not great if it was a real experiment). We can assume that each sample takes .0025 s.

# sample-1 takes the number of rows minus 1 because the first value should be 0. 
```{r}
hp$t <-  (hp$sample-1) * 0.0025
```


Part 7. Calculate velocity and acceleration for y, i.e. vertical movement. Remember: Velocity = distance difference / time difference. Acceleration = velocity difference / time difference.

```{r}
hp$vel <- c(0,diff(hp$y) / diff(hp$t))
hp$acc <- c(0,diff(hp$vel) / diff(hp$t))
plot(hp$t,hp$acc,'l')
lines(hp$t,hp$vel,'l')
lines(hp$t,hp$y,'l')
title("position, velocity, acceleration")
```

Part 8. Play around with some filter to smoothen velocity and acceleration (most common: Butterworth filter).

```{r}
bf <- butter(2, 0.1, type ='low') # requires signal package
hp$velf <- filtfilt(bf,hp$vel)
plot(hp$t,hp$vel,'l', col = "black")
lines(hp$t,hp$velf,'l', col = "red")
title(paste("velocity (filter cutoff ",0.1,")",sep=""))

bf <- butter(2, 0.01, type ='low')
velf_test <- filtfilt(bf,hp$vel)
plot(hp$t,hp$vel,'l', col = "black")
lines(hp$t,velf_test,'l', col = "red")
title(paste("velocity (filter cutoff ",0.01,")",sep=""))

bf <- butter(2, 0.01, type ='low')
hp$accf <- filtfilt(bf,hp$acc)
plot(hp$t,hp$acc,'l', col = "black")
lines(hp$t,hp$accf,'l', col = "red")
title(paste("acceleration (filter cutoff ",0.1,")",sep=""))
```

Part 9. Using zero crossings in the velocity, calculate y position extreme values (maxima and minima). Mark those extreme values on top of the x/y plot you made earlier.

Part 10. Using e.g. a range of x positions, specify a segment during which the critical movement happens, i.e. the 6th movement is what we would want to compare in this study. Mark the peak in that segment in a different color.

```{r}
# she chose to work with the first filter, therefore we use velf instead of vel. 
# Sign() gives -1 each time there is a negative value and 1 for each positive value. Buy finding the difference between each value after each other with dff(), we will get -2 each time there it shifts from being a positive to a negative velocity. And a2 when it goes from being a negative to a positive velocity. 
zcr <- c(0,diff(sign(hp$velf)))
zcrMin <- zcr==2
zcrMax <- zcr==-2
plot(hp$x,hp$y,'l')
points(hp$x[zcrMin],hp$y[zcrMin],col = "black")
points(hp$x[zcrMax],hp$y[zcrMax],col = "green")

#Maybe eye-balling the data to find the maxium of the 6th movement
zcrTarget <- hp$x>1160&hp$x<1330&zcrMax
paste("Target position (x) and value (y): ",hp$x[zcrTarget],hp$y[zcrTarget])
points(hp$x[zcrTarget],hp$y[zcrTarget],col = "red")
title(paste("participant",participant,"trial",trial,"condition",hp$cond[1]))
```
